Towards a probability measure approach for Differential Privacy in Bayesian inference

### 2018-07-18 main-sketch.tex
pass on concnlusions and futre work

\fancyhf{} % Remove fancy page headers 
\fancyhead[C]{Anonymous submission \#9999 to ACM CCS 2017} % TODO: replace 9999 with your paper number
\fancyfoot[C]{\thepage}

\setcopyright{none} % No copyright notice required for submissions
\acmConference[Anonymous Submission to ACM CCS 2017]{ACM Conference on Computer and Communications Security}{Due 19 May 2017}{Dallas, Texas}
\acmYear{2017}

@@

\begin{document}
\title{Sketch of DP - Bayesian Inference} % TODO: replace with your title

\begin{abstract}
Bayesian inference is a statistical method which allows to derive a
\emph{posterior} distribution over a parameter, starting from a
\emph{prior} distribution and observed data.  In a setting where the
data must be protected, different approaches have been taken to make
this process differentially private. \citet{dimitrakakis2014robust},
and \citet{wang2015privacy}, for instance, proved that, under specific
conditions, sampling from the posterior distribution is already
differentially private. Other works,
e.g. \cite{zhang2016differential}, \cite{foulds2016theory}, designed
differentially private mechanisms that output a representation of the
full posterior distribution. Also, accuracy of the mechanisms was
measured using metrics on the space of the numeric parameters of the
posterior distribution, e.g. $\ell_1$-norm.

@@

% TODO: replace this section with code generated by the tool at https://dl.acm.org/ccs.cfm
\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10002978.10003029.10011703</concept_id>
<concept_desc>Security and privacy~Usability in security and privacy</concept_desc>
<concept_significance>500</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc{Security and privacy~Use https://dl.acm.org/ccs.cfm to generate actual concepts section for your paper}
% -- end of section to replace with generated code

\keywords{differential privacy, Bayesian inference, Hellinger distance} % TODO: replace with your keywords

@@

% \input{main-sketch} % TODO: replace with your brilliant paper!
% \input{macros}
\section{An informal motivation}
\label{sec_intro}
Publishing the posterior distribution inferred from a sensitive dataset can
leak information about individuals on the dataset.
So, we need to noise the posterior belief in order to
protect the data set before releasing it. The question of how much noise to add
is usually answered with an $\epsilon$, and a sensitivity value.
Sensitivity can be computed in many different ways based on which metric space
is imposed on the output set of the mechanism, and yet in literature (\cite{zhang2016differential,xiao2012bayesian} )
only metrics over the numeric parameters, e.g. $\ell_1$ norm, are used.
Another approach, which we will explore here, is to impose a metric over probability measures.
An orthogonal question is that of how to measure the accuracy. Again,
that can be answered in different ways based on the metric imposed on the output space, and yet again
only in few works \todo{cite: I Think in one paper Dimitrakakis uses KL divergence to measure accuracy }
distances over probability measures have been used for these purposes.

The question that this work aims at answering is whether
an approach based on probability measures can outperform
in term of accuracy approaches based on metric over
the numeric parameters of the distributions.

We will see that in some cases this can happen.
\section{Bayesian Inference Background}
\label{sec_background}
Given a prior belief $\Pr(\theta)$ on some parameter $\theta$,
and given an observation $\dataobs$,
the posterior distribution on $\theta$ given $\dataobs$ is computed as:
\[
  \Pr(\theta | \dataobs) = \frac{\Pr(\dataobs | \theta) \cdot \Pr(\theta)}{\Pr(\dataobs)}
\]
where the expression  $\Pr(\dataobs | \theta)$ denotes the \emph{likelihood} of $\theta$ when
$\dataobs$ is observed. Since we consider $\dataobs$ to be fixed the likelihood is a function of $\theta$.
For the same reason $\Pr(\dataobs)$ is a constant independent of $\theta$.
Usually in statistics the prior distribution $\Pr(\theta)$ is chosen so that it represents
the initial belief on $\theta$. That is, when no data has been observed. In practice though,
prior distributions and likelihood functions are usually chosen so that the posterior
belongs to the same \emph{family} of distributions. In this case we say that the prior
is conjugate of the likelihood function. Using conjugate priors simplify calculcations and allows
for inference to be performed in a recursive fashion over the data.
In this work we will consider a specific instance of Bayesian inference and one of its generalizations.
Specifically, we will consider the situation where $\theta$ represents the parameter --informally called \emph{bias}--
of a Bernoulli distributed random variable, and its immediate generalization where the parameter $\vtheta$ represents
the the vector of parameters of a categorical distributed random variable.
In the former case the prior distribution over $\theta\in [0,1]$ is going to be a beta distribution, $\betad(\alpha, \beta)$, with parameters
$\alpha,\beta\in\mathbb{R}^{+}$, and with p.d.f:
\[
  \Pr(\theta)\equiv \frac{\theta^{\alpha} (1- \theta)^{\beta}}{\betaf(\alpha,\beta)}
\]
where $\betaf(\cdot,\cdot)$ is the beta function.
The data $\dataobs$ will be a sequence of $n\in\mathbb{N}$ binary values, that is $\dataobs=\langle x_1,\dots x_n \rangle, x_i\in\{0,1\}$, and the likelihood function is:
\[
  \Pr(\dataobs | \theta)\equiv \theta^{\Delta \alpha}(1-\theta)^{n - \Delta \alpha}
\]
where $\Delta \alpha = \displaystyle\sum_{i=1}^{n}x_i$.
From this it can easily be derived that the posterior distribution is:
\[
  \Pr(\theta|\dataobs)=\betad(\alpha + \Delta \alpha,\beta + n - \Delta \alpha)
\]
In the latter case the prior distribution over $\vtheta\in [0,1]^{k}$ is given by a Dirichelet distribution, $\dirichlet(\valpha)$, for $k\in\mathbb{N}$,
and $\valpha\in(\mathbf{R^{+}})^{k}$, with p.d.f:
\[
  \Pr(\vtheta)\equiv\frac{1}{\mbetaf(\valpha)}\cdot \displaystyle\prod_{i=1}^{k}{\theta_i^{\alpha_i-1}}
\]
where $\mbetaf(\cdot)$ is the generealised beta function.
The data $\dataobs$ will be a sequence of $n\in\mathbb{N}$ values coming from a universe $\datauni$, such that $\mid\datauni \mid=k$.
The likelihood function will be:
\[
  \Pr(\dataobs|\vtheta)\equiv \displaystyle\prod_{a_i\in\datauni}\theta_{i}^{\Delta \alpha_i},
\]
with $\Delta \alpha_i=\displaystyle\sum_{j=1}^{n}\iverson{x_j=a_i}$, where in $\iverson{\cdot}$ we use
Iverson bracket notation. Denoting by $\Delta\valpha$ the vector $\langle  \Delta\alpha_1,\dots \Delta\alpha_k \rangle$ the posterior distribution over $\vtheta$ turns out to be
\[
  \Pr(\vtheta|\dataobs)=\dirichlet(\valpha+\Delta \valpha). 
\]
where $+$ denotes the component wise  sum of vectors of reals.

@@

These results although very motivating, are still not enough for real world applications.
Hence, we will continue our work in the follwoing directions:
\begin{enumerate}
  \item For now, we just have a intuitive idea on the accuracy
    behavior of our mechanisms, and not a precise formula or bound on it.
    When do our mechanisms perform better than the baseline mechanism and when they don't?
    How much influence will elements in Section \ref{sec_experiment} have on the
    accuracy? Are there any other important factors we missed? These are
    all questions w.r.t. the accuracy that we are going to explore next, and in a more
    principled and formal way.
\item  Theorem \ref{thm:privacy} provides an upper bound to the  privacy loss for
  $\hexpmech$ and $\hexpmechd$ but  not necessarily a tight one.
  Indeed, experiments have shown that the actual privacy loss in the experiments can be smaller than $\epsilon$.
  This means that we could improve accuracy, by adding less noise --
  that is noise proportional to a higher value of $\epsilon$-- but still achieve $(\epsilon, \delta)$-dp.
\item The choice of the Hellinger distance might seem quite ad-hoc. Hence,
  it is worth exploring other distances over distributions. An interesting class
  of probability metrics is the family of  $f$-divergences \cite{??}.
\end{enumerate}

>>>>

\fancyhf{} % Remove fancy page headers 
% \fancyhead[C]{Anonymous submission \#9999 to ACM CCS 2017} % TODO: replace 9999 with your paper number
\fancyfoot[C]{\thepage}

\setcopyright{none} % No copyright notice required for submissions
% \acmConference[Anonymous Submission to ACM CCS 2017]{ACM Conference on Computer and Communications Security}{Due 19 May 2017}{Dallas, Texas}
% \acmYear{2017}

@@

\begin{document}
\title{Towards a probability measure approach for Differential Privacy in Bayesian inference} % TODO: replace with your title
\subtitle{Extended Abstract}
\begin{abstract}
Bayesian inference is a statistical method which allows to derive a
\emph{posterior} distribution over a parameter, starting from a
\emph{prior} distribution and observed data.  In a setting where the
data must be protected, different approaches have been taken to make
this process differentially private. \citet{dimitrakakis2014robust},
and \citet{wang2015privacy}, for instance, proved that, under specific
conditions, sampling from the posterior distribution is already
differentially private. Other works,
e.g. \cite{zhang2016differential}, \cite{foulds2016theory}, designed
differentially private mechanisms that output a representation of the
full posterior distribution. Also, accuracy of the mechanisms was
measured using metrics on the space of the numeric parameters of the
posterior distribution, e.g. $\ell_1$-norm.

@@

% TODO: replace this section with code generated by the tool at https://dl.acm.org/ccs.cfm
% \begin{CCSXML}
% <ccs2012>
% <concept>
% <concept_id>10002978.10003029.10011703</concept_id>
% <concept_desc>Security and privacy~Usability in security and privacy</concept_desc>
% <concept_significance>500</concept_significance>
% </concept>
% </ccs2012>
% \end{CCSXML}

% \ccsdesc{Security and privacy~Use https://dl.acm.org/ccs.cfm to generate actual concepts section for your paper}
% % -- end of section to replace with generated code

\keywords{Differential privacy, Bayesian inference, Hellinger distance} % TODO: replace with your keywords

@@

% \input{main-sketch} % TODO: replace with your brilliant paper!
% \input{macros}
\section{An informal motivation}
\label{sec_intro}
Publishing the posterior distribution inferred from a sensitive dataset can
leak information about individuals on the dataset.
So, we need to noise the posterior belief in order to
protect the data set before releasing it. The question of how much noise to add
is usually answered with an $\epsilon$, and a sensitivity value.
Sensitivity can be computed in many different ways based on which metric space
is imposed on the output set of the mechanism, and yet in literature (\cite{zhang2016differential,xiao2012bayesian} )
only metrics over the numeric parameters, e.g. $\ell_1$ norm, are used.
Another approach, which we will explore here, is to impose a metric over probability measures.
An orthogonal question is that of how to measure the accuracy. Again,
that can be answered in different ways based on the metric imposed on the output space, and yet again
only in few works in literature (e.g. \cite{zhang2016differential})
distances over probability measures have been used for these purposes.

The question that this work aims at answering is whether
an approach based on probability measures can outperform
in term of accuracy approaches based on metric over
the numeric parameters of the distributions. 
We will see that in some cases this can happen.
\section{Bayesian Inference Background}
\label{sec_background}
Given a prior belief $\Pr(\theta)$ on some parameter $\theta$,
and an observation $\dataobs$, the posterior distribution on $\theta$ given $\dataobs$ is computed as:
\[
  \Pr(\theta | \dataobs) = \frac{\Pr(\dataobs | \theta) \cdot \Pr(\theta)}{\Pr(\dataobs)}
\]
where the expression  $\Pr(\dataobs | \theta)$ denotes the \emph{likelihood} of $\theta$ when
$\dataobs$ is observed. Since we consider $\dataobs$ to be fixed the likelihood is a function of $\theta$.
For the same reason $\Pr(\dataobs)$ is a constant independent of $\theta$.
Usually in statistics the prior distribution $\Pr(\theta)$ is chosen so that it represents
the initial belief on $\theta$. That is, when no data has been observed. In practice though,
prior distributions and likelihood functions are usually chosen so that the posterior
belongs to the same \emph{family} of distributions. In this case we say that the prior
is conjugate of the likelihood function. Using conjugate priors simplify calculcations and allows
for inference to be performed in a recursive fashion over the data.
In this work we will consider a specific instance of Bayesian inference and one of its generalizations.
Specifically, we will consider the situation where $\theta$ represents the parameter --informally called \emph{bias}--
of a Bernoulli distributed random variable, and its immediate generalization where the parameter $\vtheta$ represents
the the vector of parameters of a categorical distributed random variable.
In the former case the prior distribution over $\theta\in [0,1]$ is going to be a beta distribution, $\betad(\alpha, \beta)$, with parameters
$\alpha,\beta\in\mathbb{R}^{+}$, and with p.d.f:
\[
  \Pr(\theta)\equiv \frac{\theta^{\alpha} (1- \theta)^{\beta}}{\betaf(\alpha,\beta)}
\]
where $\betaf(\cdot,\cdot)$ is the beta function.
The data $\dataobs$ will be a sequence of $n\in\mathbb{N}$ binary values, that is $\dataobs=\langle x_1,\dots x_n \rangle, x_i\in\{0,1\}$, and the likelihood function is:
\[
  \Pr(\dataobs | \theta)\equiv \theta^{\Delta \alpha}(1-\theta)^{n - \Delta \alpha}
\]
where $\Delta \alpha = \displaystyle\sum_{i=1}^{n}x_i$.
From this it can easily be derived that the posterior distribution is:
\[
  \Pr(\theta|\dataobs)=\betad(\alpha + \Delta \alpha,\beta + n - \Delta \alpha)
\]
In the latter case the prior distribution over $\vtheta\in [0,1]^{k}$ is given by a Dirichelet distribution, $\dirichlet(\valpha)$, for $k\in\mathbb{N}$,
and $\valpha\in(\mathbf{R^{+}})^{k}$, with p.d.f:
\[
  \Pr(\vtheta)\equiv\frac{1}{\mbetaf(\valpha)}\cdot \displaystyle\prod_{i=1}^{k}{\theta_i^{\alpha_i-1}}
\]
where $\mbetaf(\cdot)$ is the generealised beta function.
The data $\dataobs$ will be a sequence of $n\in\mathbb{N}$ values coming from a universe $\datauni$, such that $\mid\datauni \mid=k$.
The likelihood function will be:
\[
  \Pr(\dataobs|\vtheta)\equiv \displaystyle\prod_{a_i\in\datauni}\theta_{i}^{\Delta \alpha_i},
\]
with $\Delta \alpha_i=\displaystyle\sum_{j=1}^{n}\iverson{x_j=a_i}$, where in $\iverson{\cdot}$ we use
Iverson bracket notation. Denoting by $\Delta\valpha$ the vector $\langle  \Delta\alpha_1,\dots \Delta\alpha_k \rangle$ the posterior distribution over $\vtheta$ turns out to be
\[
  \Pr(\vtheta|\dataobs)=\dirichlet(\valpha+\Delta \valpha). 
\]
where $+$ denotes the component wise  sum of vectors of reals.

@@

These results although very motivating, are still not enough for real world applications.
Hence, we will continue our work in the follwoing directions:
\begin{enumerate}
  \item For now, we just have a intuitive idea on the accuracy
    behavior of our mechanisms, and not a precise formula or bound on it.
    When do our mechanisms perform better than the baseline mechanism and when they don't?
    How much influence will elements in Section \ref{sec_experiment} have on the
    accuracy? Are there any other important factors we missed? These are
    all questions w.r.t. the accuracy that we are going to explore next, and in a more
    principled and formal way.
\item  Theorem \ref{thm:privacy} provides an upper bound to the  privacy loss for
  $\hexpmech$ and $\hexpmechd$ but  not necessarily a tight one.
  Indeed, experiments have shown that the actual privacy loss in the experiments can be smaller than $\epsilon$.
  This means that we could improve accuracy, by adding less noise --
  that is noise proportional to a higher value of $\epsilon$-- but still achieve $(\epsilon, \delta)$-dp.
\item The choice of the Hellinger distance might seem quite ad-hoc. Hence,
  it is worth exploring other distances over distributions. An interesting class
  of probability metrics is the family of  $f$-divergences \todo{\cite{??}}.
\end{enumerate}
